---
title: "OSAC Selection Report Data Viz Options"
output: 
    html_document:
        css: style.css
        toc: true
        toc_depth: 2
        toc_float: true
        df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r packages, include = FALSE}
library(dkmisc)
library(tidyverse)
library(readxl)
library(janitor)
library(gt)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(ggridges)
```


```{r get_data, include = FALSE}


# Get geodata stuff

tfff_basemap <- dk_get_tfff_basemap()

# Create tfff_counties data vector to use later
tfff_counties <- tfff_basemap %>% 
     pull(name)


# Read in geocoded cities data

cities <- read_csv("data/cities.csv") %>% 
     clean_names()

# Get TFFF cities data for rural stuff later on

tfff_cities <- read_excel("data/tfff_cities.xlsx") %>% 
     clean_names() %>% 
     mutate(location = paste(city, state, sep = ", ")) %>% 
     mutate(cntyrgnnam = str_replace(cntyrgnnam, 
                                     "N Coast", 
                                     "North Coast")) %>% 
     select(location, rural) %>% 
     mutate(rural = case_when(
          rural == TRUE ~ "Rural",
          rural == FALSE ~ "Not Rural"
     ))

# Read in student data

student_data <- read_excel("data/Ford_Scholar_Data_Reviewed_Interviewed_Awarded_2014-2018.xlsx") %>% 
     clean_names() %>% 
     mutate(year = appyer) %>% 
     mutate(status_calculated = case_when(
          str_detect(status, "Interviewed - Awarded") ~ "Recipient",
          TRUE ~ "Applicant")
     ) %>% 
     mutate(interviewed = case_when(
          str_detect(status, "Not Interviewed") ~ "Not Interviewed",
          TRUE ~ "Interviewed")
     )     %>% 
     mutate(city = str_to_title(city)) %>% 
     # Add shortened year for labels
     mutate(year_label = appyer - 2000) %>% 
     mutate(county_name = str_to_title(county_name)) %>% 
     mutate(location = paste(city, st, sep = ", ")) %>% 
     left_join(cities) %>% 
     left_join(tfff_cities)

student_data_tfff_counties <- student_data %>% 
     filter(county_name %in% tfff_counties)




```

```{r look_and_feel, include = FALSE}

# Colors

tfff_dark_green <- "#265142"
tfff_light_green <- "#B5CC8E"
tfff_orange <- "#e65100"
tfff_yellow <- "#FBC02D"
tfff_blue <- "#283593"
tfff_red <- "#B71C1C"
tfff_brown <- "#51261C"
tfff_dark_gray <- "#545454"
tfff_medium_gray <- "#a8a8a8"
tfff_light_gray <- "#eeeeee"

# Themes

loadfonts()

dk_theme <- theme_ipsum_ps(base_family = "Calibri",
                           plot_title_family = "Calibri")

theme_set(theme_ipsum_ps())

dk_pie_theme <- theme_ipsum_ps() +
     theme(legend.position = "none",
           axis.text.x = element_blank(),
           axis.text.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.title.x = element_blank(),
           axis.title.y = element_blank())

```

These are all options. Anything can be tweaked, colors changed, size of things altered, etc. Also, I can get you all the charts as PDFs so you can edit them as you see fit. 

One other note: I'm using the [IBM Plex Sans](https://fonts.google.com/specimen/IBM+Plex+Sans?selection.family=IBM+Plex+Sans) font throughout here (because I like it). I couldn't remember what font you were going to use but just let me know and I can update everything.


# Cover

```{r}

student_data_tfff_counties_sf <- student_data_tfff_counties %>% 
     filter(appyer == 2018) %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     st_as_sf(coords = c("lon", "lat"),
              crs = 4326)

ggplot() +
     geom_sf(data = tfff_basemap,
             fill = tfff_light_gray) +
     geom_sf(data = filter(student_data_tfff_counties_sf, status_calculated == "Recipients"),
             # size = 1,
             shape = 21,
             fill = tfff_light_green,
             color = tfff_dark_green) +
     coord_sf(datum = NA) +
     theme(legend.position = "bottom",
           plot.title = element_text(hjust = 0.5)) +
     labs(title = "2018 Recipients")

ggsave("plots-new/2018-recipients.pdf",
       device = cairo_pdf)

```




# Region/County

## ‘Dot’ map graphic showing 2018 Applicants by county


```{r}
students_by_county_2018 <- student_data_tfff_counties %>% 
     filter(appyer == 2018) %>% 
     group_by(status_calculated) %>% 
     group_by(county_name, status_calculated) %>% 
     summarize(n = n()) %>% 
     arrange(status_calculated) %>% 
     ungroup() %>% 
     group_by(status_calculated) %>% 
     mutate(pct_from_county = prop.table(n)) %>% 
     ungroup() %>% 
     add_row(county_name = "Sherman") %>% 
     complete(county_name, status_calculated,
              fill = list(n = 0,
                          pct_from_county = 0)) %>% 
     drop_na(status_calculated) %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     right_join(tfff_basemap, by = c("county_name" = "name")) %>%
     st_as_sf(crs = 4326) 



ggplot() +
     geom_sf(data = students_by_county_2018,
             aes(fill = pct_from_county)) +
     geom_sf(data = student_data_tfff_counties_sf,
             size = 0.5,
             color = tfff_dark_green) +
     coord_sf(datum = NA) +
     scale_fill_gradient(low = "white",
                         high = tfff_dark_green,
                         labels = percent_format(1),
                          breaks = seq(0, .15, .05)) +
     # scale_fill_viridis_c(option = "C",
     #                      labels = percent_format(1),
     #                      breaks = seq(0, .15, .05)) +
     facet_wrap(~status_calculated) +
     theme(legend.position = "bottom") +
     labs(fill = "Percent of Total by County",
          title = "Recipients and Applicants by Location")

ggsave("plots-new/2018-recipients_applicants-by-location.pdf",
       device = cairo_pdf)
```



## Pie chart showing what % of 2018 recipients could be considered Rural

Figure shows percent from rural places.

```{r}
student_data %>% 
     filter(status_calculated == "Recipient") %>% 
     filter(appyer == 2018) %>% 
     count(rural) %>% 
     drop_na(rural) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     ggplot(aes(year, pct,
                fill = rural)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
     scale_fill_manual(values = c(tfff_light_gray, tfff_dark_green)) +
     dk_pie_theme

ggsave("plots-new/rural-pie.pdf",
       device = cairo_pdf)
```


## 5 OR 10 year average % of recipients considered rural

This shows data from last 5 years.

```{r}

student_data %>% 
     filter(status_calculated == "Recipient") %>% 
     count(rural) %>% 
     drop_na(rural) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = "last 5") %>% 
     ggplot(aes(year, pct,
                fill = rural)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
     scale_fill_manual(values = c(tfff_light_gray, tfff_dark_green)) +
     dk_pie_theme

ggsave("plots-new/rural-pie-5-years.pdf",
       device = cairo_pdf)
```




This shows the last 5 years as an area chart.

```{r}
student_data %>% 
     filter(status_calculated == "Recipient") %>% 
     count(appyer, rural) %>% 
     drop_na(rural) %>% 
     group_by(appyer) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     filter(rural == "Rural") %>% 
     ggplot(aes(appyer, pct)) +
     geom_area(fill = tfff_dark_green) +
     geom_line(color = tfff_dark_green) +
     scale_y_continuous(labels = percent_format(1)) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.minor.x = element_blank(),
           panel.grid.major.x = element_blank(),
           panel.grid.minor.y = element_blank())

ggsave("plots-new/rural-area-chart.pdf",
       device = cairo_pdf)
```

Same data as a line chart.
```{r}
student_data %>% 
     filter(status_calculated == "Recipient") %>% 
     count(appyer, rural) %>% 
     drop_na(rural) %>% 
     group_by(appyer) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     filter(rural == "Rural") %>% 
     ggplot(aes(appyer, pct)) +
     # geom_area(fill = tfff_dark_green) +
     geom_line(color = tfff_dark_green) +
     scale_y_continuous(labels = percent_format(1),
                        limits = c(0, .7)) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.minor.x = element_blank(),
           panel.grid.major.x = element_blank(),
           panel.grid.minor.y = element_blank())

ggsave("plots-new/rural-line-chart.pdf",
       device = cairo_pdf)
```


And this shows the last ~25 years as an area chart.

```{r}
rural_25_years <- tibble(
     year = seq(1994, 2018, 1),
     pct = c(.79, .7, .66, .71, .63, .76, .68, .73, .71, .67, .7, .67, .81, .71, .78, .73, .677, .72, .58, .66, .61, .63, .67, .63, .65)
)

ggplot(rural_25_years, aes(year, pct)) +
     geom_area(fill = tfff_dark_green) +
     geom_line(color = tfff_dark_green) +
     scale_y_continuous(labels = percent_format(1)) +
     scale_x_continuous(breaks = seq(1995, 2020, 5),
                        limits = c(1994, 2018)) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.minor.x = element_blank(),
           panel.grid.major.x = element_blank(),
           panel.grid.minor.y = element_blank())

ggsave("plots-new/rural-area-25-years.pdf",
       device = cairo_pdf)

```

Same data as a line chart.

```{r}
ggplot(rural_25_years, aes(year, pct)) +
     # geom_area(fill = tfff_dark_green) +
     geom_line(color = tfff_dark_green) +
     scale_y_continuous(labels = percent_format(1),
                        limits = c(0, .85)) +
     scale_x_continuous(breaks = seq(1995, 2020, 5),
                        limits = c(1994, 2018)) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.minor.x = element_blank(),
           panel.grid.major.x = element_blank(),
           panel.grid.minor.y = element_blank())

ggsave("plots-new/rural-line-25-years.pdf",
       device = cairo_pdf)
```



## Siskiyou county callout- What % of CA applicants become recipients VS. what % of OR applicants become recipients

Sorry about the text being small on the Oregon chart. I can adjust this later (or you can edit from the PDF).



```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     count(st, status_calculated) %>% 
     group_by(st) %>% 
     mutate(pct = prop.table(n)) %>% 
     select(st, status_calculated, pct) %>% 
     mutate(year = 2018) %>% 
     ggplot(aes(year, pct,
                fill = status_calculated)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
     scale_fill_manual(values = c(tfff_light_gray, tfff_dark_green)) +
     facet_wrap(~st) +
     labs(fill = NULL) +
     dk_pie_theme +
     theme(strip.text = element_text(hjust = 0.5))

ggsave("plots-new/siskiyou-pie.pdf",
       device = cairo_pdf)
```


## If there is one or two high schools that have more than 2 recipients- can we call them out here and possible get a quote?

Here are all high schools with 3 or more recipients in 2018.

```{r}
student_data %>%
     filter(appyer == 2018) %>%
     filter(status_calculated == "Recipient") %>%
     mutate(hs = str_glue("{hs_nam}")) %>% 
     count(hs_nam) %>%
     arrange(desc(n)) %>%
     filter(n >= 3) %>%
     set_names(c("High School", "Number of Recipients")) %>%
     gt()

```


# Gender

## Pie Chart 2018 Applicants + Recipients by gender

This shows all genders, but you may want to highlight just males since I know the board cares about that in particular.

```{r}

student_data %>% 
     filter(appyer == 2018) %>% 
     select(people_id,
            appyer,
            year_label,
            gender,
            status_calculated) %>% 
     # Need to create gender categories because there are multiple in single cells
     mutate(gender_single = case_when(
          str_detect(gender, paste(c("DI", "TF", "TM", "F, M"), collapse = "|")) ~ "Transgender/Non-Binary",
          str_detect(gender, "N") ~ "Choose not to answer",
          str_detect(gender, "M") ~ "Male",
          str_detect(gender, "F") ~ "Female"
     )) %>% 
     filter(gender_single != "Choose not to answer") %>% 
     count(status_calculated, gender_single) %>% 
     group_by(status_calculated) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     ggplot(aes(year, pct,
                fill = gender_single)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
     scale_fill_manual(values = c(tfff_light_green, tfff_dark_green, tfff_yellow)) +
     labs(fill = NULL) +
     facet_wrap(~status_calculated) +
     dk_pie_theme +
     theme(legend.position = "bottom")

ggsave("plots-new/gender-2018.pdf",
       device = cairo_pdf)
```

This chart is the same, but just highlights males.

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     select(people_id,
            appyer,
            year_label,
            gender,
            status_calculated) %>% 
     # Need to create gender categories because there are multiple in single cells
     mutate(gender_single = case_when(
          str_detect(gender, paste(c("DI", "TF", "TM", "F, M"), collapse = "|")) ~ "Transgender/Non-Binary",
          str_detect(gender, "N") ~ "Choose not to answer",
          str_detect(gender, "M") ~ "Male",
          str_detect(gender, "F") ~ "Female"
     )) %>% 
     filter(gender_single != "Choose not to answer") %>% 
     mutate(male = case_when(
          gender_single == "Male" ~ "Male",
          TRUE ~ "Not Male"
     )) %>% 
     count(status_calculated, male) %>% 
     group_by(status_calculated) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     ggplot(aes(year, pct,
                fill = male)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_gray, tfff_light_gray)) +
     labs(fill = NULL) +
     facet_wrap(~status_calculated) +
     dk_pie_theme +
     theme(legend.position = "bottom")

ggsave("plots-new/gender-2018-males-only.pdf",
       device = cairo_pdf)
```



## Stacked Column bar chart showing gender over the last 10 years

I only have the data for males. I also only have the 2017 Excel spreadsheet so this shows data from 1994-2017.

```{r}


tibble::tribble(
     ~year,   ~pct,
     1994, "28%",
     1995, "35%",
     1996, "32%",
     1997, "44%",
     1998, "41%",
     1999, "33%",
     2000, "47%",
     2001, "35%",
     2002, "30%",
     2003, "37%",
     2004, "38%",
     2005, "41%",
     2006, "38%",
     2007, "35%",
     2008, "31%",
     2009, "41%",
     2010, "38%",
     2011, "29%",
     2012, "37%",
     2013, "31%",
     2014, "33%",
     2015, "30%",
     2016, "28%",
     2017, "40%"
) %>% 
     mutate(pct = parse_number(pct)) %>% 
     mutate(pct = pct / 100) %>% 
     ggplot(aes(year, pct)) +
     geom_area(fill = tfff_dark_green) +
     geom_line(color = tfff_dark_green) +
     scale_y_continuous(labels = percent_format(1)) +
     scale_x_continuous(breaks = seq(1995, 2020, 5),
                        limits = c(1994, 2017)) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.minor.x = element_blank(),
           panel.grid.major.x = element_blank(),
           panel.grid.minor.y = element_blank())


ggsave("plots-new/males-25ish-years.pdf",
       device = cairo_pdf)


```



# Race/Ethnicity

## Pie chart showing Ethnicity of 2018 applicants + interviewees + recipients

This shows the percent non-white for each group.

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     select(status, ethnic_group) %>% 
     mutate(race_ethnicity = case_when(
          ethnic_group == "WHITE" ~ "White",
          TRUE ~ "Non-White"
     )) %>% 
     count(status, race_ethnicity) %>% 
     group_by(status) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     mutate(status = case_when(
          status == "Interviewed - Awarded" ~ "Recipients",
          status == "Interviewed - Not Awarded" ~ "Interviewees",
          status == "Reviewed - Not Interviewed" ~ "Applicants"
     )) %>% 
     ggplot(aes(year, pct,
                fill = race_ethnicity)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_gray)) +
     labs(fill = NULL) +
     facet_wrap(~status) +
     dk_pie_theme +
     theme(strip.text = element_text(hjust = 0.5),
           plot.title = element_text(hjust = 0.5))

ggsave("plots-new/race-2018-pies.pdf",
       device = cairo_pdf)
```



## Line Graph comparing latino recipients over the last 10 years compared to the STATE Latino population (if data is available)

It's proving a bit more challenging to get the state data than I had hoped. If you really want to include it, I can keep digging. If not, let's leave it for now.

```{r}
tibble::tribble(
     ~year,  ~pct,
     1994, 0.19,
     1995, 0.04,
     1996, 0.06,
     1997, 0.05,
     1998, 0.04,
     1999, 0.12,
     2000, 0.07,
     2001, 0.05,
     2002, 0.13,
     2003, 0.09,
     2004, 0.09,
     2005,  0.1,
     2006, 0.12,
     2007, 0.13,
     2008, 0.13,
     2009, 0.19,
     2010,  0.2,
     2011, 0.23,
     2012, 0.11,
     2013, 0.25,
     2014,  0.2,
     2015, 0.31,
     2016, 0.24,
     2017, 0.43
) %>% 
     ggplot(aes(year, pct)) +
     # geom_area(fill = tfff_dark_green) +
     geom_line(color = tfff_dark_green) +
     scale_y_continuous(labels = percent_format(1)) +
     scale_x_continuous(breaks = seq(1995, 2020, 5),
                        limits = c(1994, 2017)) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.minor.x = element_blank(),
           panel.grid.major.x = element_blank(),
           panel.grid.minor.y = element_blank())

ggsave("plots-new/latinos-25ish-years.pdf",
       device = cairo_pdf)

```



# First Generation/EFC

## Percentage of 2018 applicants + recipients who are First Generation

Figure shows % of students who are first gen.

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     select(status_calculated, mothered, fathered) %>% 
     mutate(first_gen = case_when(
          mothered != 3 & fathered != 3 ~ "First Generation",
          TRUE ~ "Not First Generation"
     )) %>% 
     count(status_calculated, first_gen) %>% 
     group_by(status_calculated) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     ggplot(aes(year, pct,
                fill = first_gen)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_gray)) +
     labs(fill = NULL) +
     facet_wrap(~status_calculated) +
     dk_pie_theme +
     theme(strip.text = element_text(hjust = 0.5),
           plot.title = element_text(hjust = 0.5))

ggsave("plots-new/first-gen-2018-pies.pdf",
       device = cairo_pdf)
```


## Statement about 10 year average of first gen recipients

We
```{r}
student_data %>% 
     select(status_calculated, mothered, fathered) %>% 
     mutate(first_gen = case_when(
          mothered != 3 & fathered != 3 ~ "First Generation",
          TRUE ~ "Not First Generation"
     )) %>% 
     count(status_calculated, first_gen) %>% 
     group_by(status_calculated) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = "Last 5") %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     ggplot(aes(year, pct,
                fill = first_gen)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)), 
               position = position_stack(vjust = 0.5),
               color = tfff_light_gray) +
     coord_polar("y", direction = -1) +
      scale_fill_manual(values = c(tfff_dark_green, tfff_light_gray)) +
     labs(fill = NULL) +
     facet_wrap(~status_calculated) +
     dk_pie_theme +
     theme(strip.text = element_text(hjust = 0.5),
           plot.title = element_text(hjust = 0.5))

ggsave("plots-new/first-gen-10-years.pdf",
       device = cairo_pdf)
```

## Table on Expected Family Contribution

This is only for 2018. I showed this to two decimal places (otherwise several items round to 0 percent), though I'd not suggest doing this in the final report because I think this is more detail than people need. I'd put anything super small as "less than 1%".

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     tabyl(efclvl, status_calculated) %>% 
     adorn_percentages("col") %>% 
     adorn_pct_formatting(digits = 2, rounding = "half up") %>% 
     adorn_ns() %>% 
     mutate(efcmin = (parse_number(efclvl) - 1) * 1000) %>% 
     mutate(efcmax = efcmin + 1000) %>% 
     mutate(efc = str_glue("${efcmin}-${efcmax}")) %>% 
     select(efc, Applicant, Recipient) %>% 
     rename("EFC Level" = "efc") %>% 
     gt()
```


# Back Page

## Rural

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     count(rural, status_calculated) %>% 
     drop_na(rural) %>% 
     group_by(status_calculated) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     filter(rural == "Rural") %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     ggplot(aes(status_calculated, pct,
                fill = status_calculated)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)),
               vjust = 2,
               color = "white") +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_green)) +
     labs(fill = NULL) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.y = element_blank(),
           legend.position = "none")

ggsave("plots-new/rural-2018-bar.pdf",
       device = cairo_pdf)
```

## Gender

This chart just shows % of males.

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     mutate(gender_single = case_when(
          str_detect(gender, paste(c("DI", "TF", "TM", "F, M"), collapse = "|")) ~ "Transgender/Non-Binary",
          str_detect(gender, "N") ~ "Choose not to answer",
          str_detect(gender, "M") ~ "Male",
          str_detect(gender, "F") ~ "Female"
     )) %>% 
     filter(gender_single != "Choose not to answer") %>% 
     count(gender_single, status_calculated) %>% 
     drop_na(gender_single) %>% 
     group_by(status_calculated) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     filter(gender_single == "Male") %>%
     ggplot(aes(status_calculated, pct,
                fill = status_calculated)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)),
               vjust = 2,
               color = "white") +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_green)) +
     labs(fill = NULL) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.y = element_blank(),
           legend.position = "none")

ggsave("plots-new/males-2018-bar.pdf",
       device = cairo_pdf)
```

## First Generation

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     mutate(first_gen = case_when(
          mothered != 3 & fathered != 3 ~ "First Generation",
          TRUE ~ "Not First Generation"
     )) %>% 
     count(first_gen, status_calculated) %>% 
     drop_na(first_gen) %>% 
     group_by(status_calculated) %>% 
     mutate(pct = prop.table(n)) %>% 
     ungroup() %>% 
     mutate(year = 2018) %>% 
     filter(first_gen == "First Generation") %>%
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     ggplot(aes(status_calculated, pct,
                fill = status_calculated)) +
     geom_col() +
     geom_text(aes(label = percent(pct, 1)),
               vjust = 2,
               color = "white") +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_green)) +
     labs(fill = NULL) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.y = element_blank(),
           legend.position = "none")

ggsave("plots-new/first-gen-2018-bar.pdf",
       device = cairo_pdf)
```

## EFC Contribution

The data for this isn't exact numbers (it's, for example, $0-$1000) so we need to figure out how to show this. It could be, say, $2000 or less. 

## Average GPA & SAT (maybe this is in one table)

### Math SAT

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     group_by(status_calculated) %>% 
     summarize(math_sat = mean(satmth, na.rm = TRUE)) %>% 
     mutate(year = 2018) %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     ggplot(aes(status_calculated, math_sat,
                fill = status_calculated)) +
     geom_col() +
     geom_text(aes(label = number(math_sat, 1)),
               vjust = 2,
               color = "white") +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_green)) +
     labs(fill = NULL) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.y = element_blank(),
           legend.position = "none")

ggsave("plots-new/math-sat-2018-bar.pdf",
       device = cairo_pdf)
 
```

### Verbal SAT

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     group_by(status_calculated) %>% 
     summarize(verbal_sat = mean(satmth, na.rm = TRUE)) %>% 
     mutate(year = 2018) %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     ggplot(aes(status_calculated, verbal_sat,
                fill = status_calculated)) +
     geom_col() +
     geom_text(aes(label = number(verbal_sat, 1)),
               vjust = 2,
               color = "white") +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_green)) +
     labs(fill = NULL) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.y = element_blank(),
           legend.position = "none")


ggsave("plots-new/verbal-sat-2018-bar.pdf",
       device = cairo_pdf)
 
 
```


### HS GPA

```{r}
student_data %>% 
     filter(appyer == 2018) %>% 
     group_by(status_calculated) %>% 
     summarize(gpa = mean(hsgpa, na.rm = TRUE)) %>% 
     mutate(year = 2018) %>% 
     mutate(status_calculated = str_glue("{status_calculated}s")) %>% 
     ggplot(aes(status_calculated, gpa,
                fill = status_calculated)) +
     geom_col() +
     geom_text(aes(label = number(gpa, .01)),
               vjust = 2,
               color = "white") +
     scale_fill_manual(values = c(tfff_dark_green, tfff_light_green)) +
     labs(fill = NULL) +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.y = element_blank(),
           legend.position = "none")


ggsave("plots-new/hs-gpa-2018-bar.pdf",
       device = cairo_pdf)
 
 
```

